Exploring how changes in age at death and cause of death have narrowed the black-white life expectancy gap over forty-five years in the United States.

### 2017-02-06 Manuscript/Manuscript_using_actual_data.Rmd
graphs by census division

age_cod_results <-rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_1.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_2.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_3.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_4.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_5.csv"))

age_cod_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_1.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_2.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_3.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_4.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_5.csv"))

age_decomp_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_1.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_2.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_3.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_4.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_5.csv"))

BlackWhite_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_1.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_2.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_3.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_4.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_5.csv"))

cod_decomp_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_1.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_2.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_3.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_4.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_5.csv"))
```

@@

```{r add_Census_Region, echo=FALSE}
BlackWhite_results$Census_Region <- "South"
BlackWhite_results$Census_Region[BlackWhite_results$state %in% c("Iowa", "Kansas", "Minnesota", "Missouri", 
                                                 "Nebraska", "North Dakota", "South Dakota", 
                                                 "Illinois", "Indiana", "Michigan", "Ohio", 
                                                 "Wisconsin")] <- "Midwest"

@@

```{r addvar_State_abbreviation, echo=FALSE}
BlackWhite_results <- BlackWhite_results %>% mutate(stabbrs = factor(state,  
                                                                     labels = c("AL", "AZ", "AR", "CA",
                                                                                "CO", "CT", "DE",
                                                                                "FL", "GA", "IL", "IN",
                                                                                "IA", "KS", "KY", "LA",
                                                                                "MD", "NY", "NC", "TX",
                                                                                "VA", "DC", "WI")))

@@

*Changes to the differency in life expectancy in US males*

@@

*Changes to the differency in life expectancy in US females*

@@

```{r figure5a, fig.align="center", fig.height=4, fig.width=11, echo=F}
ggplotly(ggplot(subset(BlackWhite_results, sex == "Male"), aes(y=LE_white, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_ribbon(aes(ymin = LE_white_lcl, ymax = LE_white_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) +
           facet_grid(. ~ Census_Region) +
           ylab("Life expectancy (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite_results, year == 2013 & sex == "Male"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))

@@

```{r figure5b, fig.align="center", fig.height=4, fig.width=11, echo=F}
ggplotly(ggplot(subset(BlackWhite_results, sex == "Male"), aes(y=LE_black, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_ribbon(aes(ymin = LE_black_lcl, ymax = LE_black_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) +
           facet_grid(. ~ Census_Region) +
           ylab("Life expectancy (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite_results, year == 2013 & sex == "Male"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))

@@

```{r figure5c, fig.align="center", fig.height=4, fig.width=11, echo=F}
ggplotly(ggplot(subset(BlackWhite_results, sex == "Male"), aes(y=LE_wbgap, x = year)) + 
           geom_vline(aes(xintercept = 1983), lty = 2, lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), lty = 2, lwd = 0.5) +
           geom_ribbon(aes(ymin = LE_wbgap_lcl, ymax = LE_wbgap_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) +
           facet_grid(. ~ Census_Region) +
           ylab("Black-white life expectancy gap (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite_results, year == 2013 & sex == "Male"), aes(label = stabbrs), check_overlap = T, size = 2.5))

@@

**Figure 6a: Contribution of cancer to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**
```{r figure6a, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cancers"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(.~ Census_Region) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cancers"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 6a (ii): Contribution of cancer (%) to the black-white life expectancy gap for males between 1969 and 2013 by US state**
```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cancers"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(.~ Census_Region) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cancers"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cardiovascular"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(.~ Census_Region) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cardiovascular"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cardiovascular"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cardiovascular"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Communicable"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Communicable"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Non-communicable"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Non-communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Non-communicable"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Non-communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Injuries"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Injuries"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 6f: Contribution of all other causes to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**
```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "All other causes"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_grid(. ~ Census_Region) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "All other causes"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

>>>>

age_cod_results <-rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_1.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_2.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_3.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_4.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_5.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_6.csv"),
                        read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_7.csv")#,
                        #read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_8.csv")
                        )

age_cod_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_1.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_2.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_3.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_4.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_5.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_6.csv"),
                         read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_7.csv")#,
                         #read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_set_8.csv")
                         )

age_decomp_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_1.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_2.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_3.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_4.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_5.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_6.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_7.csv")#,
                            #read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_set_8.csv")
                            )

BlackWhite_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_1.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_2.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_3.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_4.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_5.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_6.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_7.csv")#,
                            #read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_set_8.csv")
                            )

cod_decomp_results <- rbind(read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_1.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_2.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_3.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_4.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_5.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_6.csv"),
                            read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_7.csv")#,
                            #read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_marginal_set_8.csv")
                            )
```

@@

```{r add_Census_Region, echo=FALSE}
#note that the Census Region of the South is partitioned into two for better display.

BlackWhite_results$Census_Region <- "South"
BlackWhite_results$Census_Region[BlackWhite_results$state %in% c("Iowa", "Kansas", "Minnesota", "Missouri", 
                                                 "Nebraska", "North Dakota", "South Dakota", 
                                                 "Illinois", "Indiana", "Michigan", "Ohio", 
                                                 "Wisconsin")] <- "Midwest"

@@

```{r add_Census_Division, ncol = 3, echo = F}
BlackWhite_results$Census_Division <- NA

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Connecticut", "Maine", "Massachusetts", 
                                                                   "New Hampshire", "Rhode Island", "Vermont")] <- "New England"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("New Jersey", "New York", "Pennsylvania")] <- "Mid-Atlantic"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Illinois", "Indiana", "Michigan", "Ohio", 
                                                                 "Wisconsin")] <- "East North Central"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Iowa", "Kansas", "Minnesota", "Missouri", 
                                                                 "Nebraska", "North Dakota", "South Dakota")] <- "West North Central"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Delaware", "Florida", "Georgia", "Maryland", 
                                                                 "North Carolina", "South Carolina", 
                                                                 "Virginia", "Washington DC", "West Virginia")] <- "South Atlantic"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Alabama", "Kentucky", "Mississippi", "Tennessee")] <- "East South Central"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Arkansas", "Louisiana", "Oklahoma", "Texas")] <- "West South Central"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Arizona", "Colorado", "Idaho", "Montana", 
                                                                 "Nevada", "New Mexico", "Utah", "Wyoming")] <- "Mountain"

BlackWhite_results$Census_Division[BlackWhite_results$state %in% c("Alaska", "California", "Hawaii", "Oregon","Washington")] <- "Pacific"

BlackWhite_results$Census_Division <- factor(BlackWhite_results$Census_Division, levels = c("New England", "Mid-Atlantic", "East North Central", "West North Central", "South Atlantic", "East South Central", "West South Central", "Mountain", "Pacific"))

```{r addvar_State_abbreviation, echo=FALSE}
BlackWhite_results <- BlackWhite_results %>% mutate(stabbrs = factor(state,  
                                                                     labels = c("AL", "AZ", "AR", "CA",
                                                                                "CO", "CT", "DE", "FL",
                                                                                "GA", "IL", "IN", "IA",
                                                                                "KS", "KY", "LA", "MD",
                                                                                "MA", "MI", "MN", "MS",
                                                                                "MO", "NE", "NV", "NJ", 
                                                                                "NM", "NY", "NC", "OH", 
                                                                                "OK", "OR", "TX", "VA", 
                                                                                "DC", "WI")))
#set 8:
#"PA", "RI", "SC", "TN", "WA", "WV"

@@

*Changes to the difference in life expectancy in US males*

@@

*Changes to the difference in life expectancy in US females*

@@

```{r figure5a, fig.align="center", fig.height=4, fig.width=11, echo=F}
ggplotly(ggplot(subset(BlackWhite_results, sex == "Male"), aes(y=LE_white, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_ribbon(aes(ymin = LE_white_lcl, ymax = LE_white_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Life expectancy (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite_results, year == 2013 & sex == "Male"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))

@@

```{r figure5b, fig.align="center", fig.height=4, fig.width=11, echo=F}
ggplotly(ggplot(subset(BlackWhite_results, sex == "Male"), aes(y=LE_black, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_ribbon(aes(ymin = LE_black_lcl, ymax = LE_black_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Life expectancy (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite_results, year == 2013 & sex == "Male"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))

@@

```{r figure5c, fig.align="center", fig.height=4, fig.width=11, echo=F}
ggplotly(ggplot(subset(BlackWhite_results, sex == "Male"), aes(y=LE_wbgap, x = year)) + 
           geom_vline(aes(xintercept = 1983), lty = 2, lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), lty = 2, lwd = 0.5) +
           geom_ribbon(aes(ymin = LE_wbgap_lcl, ymax = LE_wbgap_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Black-white life expectancy gap (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite_results, year == 2013 & sex == "Male"), aes(label = stabbrs), check_overlap = T, size = 2.5))

@@

**Figure 6a: Contribution of cancer to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**
```{r figure6a, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cancers"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap(~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cancers"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 6a (ii): Contribution of cancer (%) to the black-white life expectancy gap for males between 1969 and 2013 by US state**
```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cancers"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap(~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cancers"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cardiovascular"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap(~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cardiovascular"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Cardiovascular"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Cardiovascular"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Communicable"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Communicable"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Non-communicable"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Non-communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Non-communicable"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Non-communicable"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Injuries"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

@@

```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "Injuries"), aes(x = year, y = COD_cont_prop)) + 
           geom_ribbon(aes(ymin = COD_cont_prop_lcl, ymax = COD_cont_prop_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (%)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 6f: Contribution of all other causes to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**
```{r, fig.align="center", fig.height=5, fig.width=11, echo=F}
ggplotly(ggplot(subset(cod_decomp_results, sex == "Male" & COD == "All other causes"), aes(x = year, y = COD_cont_yrs)) + 
           geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl, fill = state), alpha=0.25) +
           geom_line(aes(col = state)) + facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Contribution to LE Gap (years)") +
           xlab("Year") +
           geom_text(data = subset(cod_decomp_results, year == 2013 & sex == "Male" & COD == "All other causes"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```
